---
title: "HW 1"
subtitle: "Data 180, Professor Kennedy"
author: 
  name: "Peter Guma"
  email: "gumap@dickinson.edu"
date: '`r format(Sys.Date(), "%B %d, %Y")`'
output: 
  html_document
---
Make sure that you have accepted the hw invitation link, and working with your clone repository.

**Due date:** September 21 beginning of class.

**Reminder:** You can insert R code in your Rmd file as a chunk in the following way:
```{r}
# Your code goes here
```
Make sure to insert a chunk for each (sub)section of the assignment.

## Preliminary Task
Go to [datacamp.com](https://datacamp.com) and complete the tutorial "Introduction to R".


## Task 1
a) Store the data below in a data frame object named `my_df`.


|Person | Preference|
|:------|:----------|
|Alice | Twitter |
| Bob | Facebook |
| Charlie | Twitter |
| Denise | Instagram |
| Erin | Facebook |
| Frank | Twitter |
| Gisele | Twitter |
| Harry | Instagram |

```{r}
person_vector <- c("Alice", "Bob", "Charlie", "Denise", "Erin", "Frank", "Gisele", "Harry")
preference_vector <- c("Twitter", "Facebook", "Twitter", "Instagram", "Facebook", "Twitter", "Twitter", "Instagram")
# factor_preference_vector <- factor(preference_vector)
# levels(factor_preference_vector)
my_df <- data.frame(person_vector, preference_vector)
# my_df
```

b) What are the dimensions of `my_df`?
```{r}
# str(my_df)
dim(my_df)
```
my_df is an 8x2 dataframe. That is, it contains 8 objects of 2 variables, or simply 8 rows and 2 columns.

c) Write R code that counts the total number of users.
```{r}
nrow(my_df) # number of users is equivalent to number of rows
```
There are 8 total users according to this function.

d) Write R code that counts the total number of users who prefer Twitter.
```{r}
# pref_col <- my_df[,2]
num_prefer_twitter <- nrow(subset(my_df, my_df[2] == "Twitter"))
num_prefer_twitter
# twitter_only <- my_df[1][my_df[2] == "Twitter"]
# twitter_only
```
4 users prefer Twitter according to the code above.

e) Write R code that returns what Gisele prefers.
```{r}
my_df[7, 2]
```
Gisele prefers Twitter.

## Task 2
a) Create a vector named `my_vec` that includes all integers from 1 to 10 (inclusive).

b) Replace the second element in `my_vec` with the number `99`.

c) Write a function called `my_func` that prints `This element is greater than 5!` if an element in `myvec` is greater than 5, and `This element is NOT greater than 5!` otherwise.

## Task 3
a. Write a function named `get_fibonacci` that returns the n'th value in the [fibonacci sequence](https://www.mathsisfun.com/numbers/fibonacci-sequence.html). For example, `get_fibonacci(5)` should return 3; `get_fibonacci(7)` should return 8 and so on. ***Hint:** It is possible to call the same function from inside it. E.g., you can call `get_fibonacci()` from inside `get_fibonacci()`. Like inception!*

b. Using your function in part (a), find and return the 19th value in the fibonacci sequence.


<br>
You are done! 🏁 Don't forget to `commit` and `push` your .Rmd file to your Github clone repository before the due date.
<br>

