---
title: "Data 180 Final"
subtitle: "Data 180, Professor Kennedy"
author: 
  name: "Peter Guma"
  email: "gumap@dickinson.edu"
date: '`r format(Sys.Date(), "%B %d, %Y")`'
output: 
  html_document
editor_options: 
  chunk_output_type: console
---

```{r echo=FALSE}
# Custom options for knitting
knitr::opts_chunk$set(
  message = FALSE,
  warning = FALSE,
  error = FALSE, 
  fig.align = "center",
  cache = FALSE
) 
```

```{r}
data <- read.csv("loan_default_data_set.csv")
```

1. Data wrangling:

a. What is the dimension (shape) of the dataset?  How many rows and columns does the data set have?
```{r}
dim(data)
```
The dimension of the dataset is 20000x21. That is, it has 20000 rows and 21 columns.

b. Report the column names of the data set.
The column names are tot_balance, avg_bal_cards, credit_age, credit_age_good_account, credit_card_age, num_acc_30d_past_due_12_months, num_acc_30d_past_due_6_months, num_mortgage_currently_past_due, tot_amount_currently_past_due, num_inq_12_month, num_card_inq_24_month, num_card_12_month, num_auto_36_month, uti_open_card, pct_over_50_uti, uti_max_credit_line, pct_card_over_50_uti, ind_XYZ, rep_income, rep_education, and Def_ind.

c. Which types of data are there in the dataset? Numeric, categorical, ordinal?
The dataset primarily consists of quantitative (i.e., numeric) data, including everything associated with the variable tot_balance. However, Def_Ind is an nominal variable: either the user did or did not default, and there is no natural ordering. While we may be tempted to label rep_education as an ordinal variable because there are different levels of education, the "other" category has no clear place in that system. I would argue that the dataset contains only numeric and nominal categorical data.

d. Which columns contain missing values and how much (what percent) of those columns are missing?
```{r}

```

e. How do you think we should deal with missing values? 

f. With this data, would you fit a supervised or an unsupervised learning model? Why? 

g. For part 2 and 3 drop all rows of the data that contain missing values. Print the dimensions of the resulting data set that has no missing values.
```{r}

```

2. Data summary statistics:

a. Find the summary statistics of the data set. You can use the summary function from dplyr. 
```{r}

```

b. Based on the mean, mode, and median, is “num_card_inq_24_month” bell shaped, left, right skewed? How about “tot_amount_currently_past_due”? “credit_age”? 
```{r}

```

c. Plot a histogram of the variables in b above. Do the shapes of the histograms confirm the skewness you found in b?
```{r}

```

d. How would your convert the “rep_education” column into numerical data? Name two ways. 


3. Data Visualization:
For every graph in this section, remember to label your axes and to include a title. Feel free to play around with graphics and parameters. Have fun and explore!

a. Plot a bar graph for the “Def_Ind” column and describe it. 
```{r}

```

b. Plot a bar graph for the “rep_education" column and describe it. 
```{r}

```

c. Plot a histogram of the “rep_income” variable.
```{r}

```

d. Plot a boxplot of the “tot_balance” variable. Using the box plot report the five number summary of the variable? Are there any outliers for this variable?
```{r}

```
