---
title: "Data 180 Final"
subtitle: "Data 180, Professor Kennedy"
author: 
  name: "Peter Guma"
  email: "gumap@dickinson.edu"
date: '`r format(Sys.Date(), "%B %d, %Y")`'
output: 
  html_document
editor_options: 
  chunk_output_type: console
---

```{r echo=FALSE}
# Custom options for knitting
knitr::opts_chunk$set(
  message = FALSE,
  warning = FALSE,
  error = FALSE, 
  fig.align = "center",
  cache = FALSE
) 
```

```{r}
data <- read.csv("loan_default_data_set.csv")
```

1. Data wrangling:

a. What is the dimension (shape) of the dataset?  How many rows and columns does the data set have?
```{r}
dim(data)
```
The dimensions of the dataset are 20000x21. That is, it has 20000 rows and 21 columns.

b. Report the column names of the data set.

The column names are tot_balance, avg_bal_cards, credit_age, credit_age_good_account, credit_card_age, num_acc_30d_past_due_12_months, num_acc_30d_past_due_6_months, num_mortgage_currently_past_due, tot_amount_currently_past_due, num_inq_12_month, num_card_inq_24_month, num_card_12_month, num_auto_36_month, uti_open_card, pct_over_50_uti, uti_max_credit_line, pct_card_over_50_uti, ind_XYZ, rep_income, rep_education, and Def_ind.

c. Which types of data are there in the dataset? Numeric, categorical, ordinal?

The dataset consists primarily of quantitative (i.e., numeric) data, including all the information associated with the variable tot_balance. However, Def_Ind is an nominal variable: either the user did or did not default, and there is no natural ordering. While we may be tempted to label rep_education as an ordinal variable because there are increasing levels of education, the "other" category has no clear place in that system. I would argue that the dataset contains only numeric and nominal categorical data.

d. Which columns contain missing values and how much (what percent) of those columns are missing?
```{r}
# is.na(data) #hidden for knit
sum(is.na(data$pct_card_over_50_uti))/20000
sum(is.na(data$rep_income))/20000
#nrow(is.na(data$pct_card_over_50_uti[data$pct_card_over_50_uti == TRUE]))
#length(data[data$pct_card_over_50_uti==TRUE])
```
The only columns which contain missing values are pct_card_over_50_uti and rep_income. The former consists of $9.79\%$ missing values, while the latter column is $7.795\%$ missing.

e. How do you think we should deal with missing values? 
One approach to dealing with missing values is to remove those rows altogether. The problem is that doing so means throwing away customer data, which is inadvisable. Removing the columns would be even worse, affecting every customer's record. Alternatively, we could replace every missing value with a placeholder like NA or 0, but that could impede data analysis. Removing the affected rows is probably our best bet in this scenario.

f. With this data, would you fit a supervised or an unsupervised learning model? Why? 

I think an unsupervised learning model would be better here because that could help us understand the structure of our multivariable dataset. If we were to conduct supervised learning, we would need to have a specific goal in mind; for example, you could make the case for rep_education predicting rep_income, or vice versa. What we get out of such a model depends on how we designate our variables (predictor or response). Unsupervised learning lacks this drawback. Applying the unsupervised technique of clustering could reveal a strong correlation between two variables that we would otherwise not think of.

g. For part 2 and 3 drop all rows of the data that contain missing values. Print the dimensions of the resulting data set that has no missing values.
```{r}
refined_data <- na.omit(data)
dim(refined_data)
```
The new data set without missing values has dimensions 16653x21.

2. Data summary statistics:

a. Find the summary statistics of the data set. You can use the summary function from dplyr. 
```{r}

```

b. Based on the mean, mode, and median, is “num_card_inq_24_month” bell shaped, left, right skewed? How about “tot_amount_currently_past_due”? “credit_age”? 
```{r}

```

c. Plot a histogram of the variables in b above. Do the shapes of the histograms confirm the skewness you found in b?
```{r}

```

d. How would your convert the “rep_education” column into numerical data? Name two ways. 


3. Data Visualization:
For every graph in this section, remember to label your axes and to include a title. Feel free to play around with graphics and parameters. Have fun and explore!

a. Plot a bar graph for the “Def_Ind” column and describe it. 
```{r}

```

b. Plot a bar graph for the “rep_education" column and describe it. 
```{r}

```

c. Plot a histogram of the “rep_income” variable.
```{r}

```

d. Plot a boxplot of the “tot_balance” variable. Using the box plot report the five number summary of the variable? Are there any outliers for this variable?
```{r}

```
