---
title: "HW 2: Visualization, Base R"
subtitle: "Data 180, Professor Kennedy"
author: 
  name: "Peter Guma"
  email: "gumap@dickinson.edu"
date: '`r format(Sys.Date(), "%B %d, %Y")`' # no need to change date since this line grabs today's date
output: 
  html_document
---

Make sure that you have accepted the hw invitation link, and working with your clone repository.

**Due date:** September 28 beginning of class.

**Reminder:** You can insert R code in your Rmd file as a chunk in the following way:
```{r}
# Your code goes here
```
Make sure to insert a chunk for each (sub)section of the assignment.

**Theme of the assignment**: Let's practice visualization using base R and create cool maps. Load in the `Cars93` dataset by running the following code chunk:

```{r}
library(MASS)
data(Cars93)
```

The library function will load and attach add-on packages that have been previously installed on your machine. This will give you access to the functions and datasets available in the package. In general, you only need to install a package once (although it may require updating from time to time) but you need to access it using the library command every time you start a new session in R (and want the functionality). 

The questions below use the `Cars93` data frame as a source of data. This data frame is part of the MASS package and becomes available after loading it. You can get a view of the top of the data frame by using the head command. For more information, type `?Cars93`. Let's load `Cars93` in,
```{r}
head(Cars93)
```

# Base R
## Task 1
a) Create a Frequency table and Relative Frequency table (latter uses percentages) for the variable `Type`. Make sure to round all digits to 3 digits.
```{r}
table(Cars93$Type) # frequency
round(table(Cars93$Type)/nrow(Cars93), 3) # relative frequency
```
b) Create a frequency bar graph for the variable `Type`. You can have count on the y axis.
```{r}
barplot(table(Cars93$Type), ylab="Count", cex.lab=1.2, cex.names=1.2, cex.axis=1.2, col=c(2,3,4,5,6,7), xlab = "Type of Car", main="Frequency by Type of 93 Cars on Sale in the USA in 1993", space=1.8)
```

*	Your axes should be labeled and legible
*	The bars should be of different colors
*	The plot should include a title using the main argument
*	The names, labels, and axes sizes (cex parameter) should be set to 1.2

c) Create a pie chart of the variable `Type`.
```{r}
pie(table(Cars93$Type), main="Frequency by Type of 93 Cars on Sale in the USA in 1993", cex=1.2)
```

* The plot should include a title using the main argument.
*	The names, labels, and axes sizes (cex parameter) should be set to 1.2

d) What is the mode of the variable `Type`? Comment.

The mode of the variable `Type` is "Midsize". That is, that type of car shows up most frequently in the dataset.

## Task 2 
a) Create a histogram of the variable `Horsepower`.
```{r}
hist(Cars93$Horsepower, col=2, xlab="Horsepower (maximum)", cex.lab=1.2, cex.axis=1.2, main="Frequency by Horsepower of 93 Cars on Sale in the USA in 1993")
# cex.names came up as being a non-graphical parameter, so I couldn't specify it
```

*	Your axes should be labeled
*	The hist body should have a color (your choice, not too dark to obscure lines) 
*	The plot should include a title using the main argument
*	The names, labels, and axes sizes should be set to 1.2

b) Create a second histogram with breaks at 25 HP increments instead of 50.
```{r}
hist(Cars93$Horsepower, col=2, xlab="Horsepower (maximum)", cex.lab=1.2, cex.axis=1.2, main="Frequency by Horsepower of 93 Cars on Sale in the USA in 1993", breaks=25)
```

* Use the breaks argument to define the classes
* Other setting should be as in part (a)

c) Which histogram more closely follows or rule of thumb for the suggested number of classes? Explain.

The second histogram more closely follows our rule of thumb for the suggested number of classes because 25 is closer than five to √n, where n is the range (max-min). In this case, √n is about 15.8. Our audience would miss some crucial information about how the values are distributed with only five classes, namely that four or five ten-horsepower intervals contain most of the cars. While five classes may be more visually appealing, 25 is more informative.

## Task 3 
a) What is the type of variable `Passengers`? (Categorical nominal/ordinal or Numerical discrete/continuous?) What happens if we were to make a histogram for `Passengers`? What is the problem with this graph?
```{r}
class(Cars93$Passengers)
hist(Cars93$Passengers)
```
The variable `Passengers` is numerical discrete. If we were to make a histogram for `Passengers`, every data point would be on the border between classes because they are all positive integers. Histograms are meant for continuous data, not discrete. It would be very difficult to tell, for instance, what is causing the class between 4 and 5 passengers to be so large. The mode is hence ambiguous.

b) Correct the problem in part (a) by creating a more appropriate visualization. Make sure to label your axes, place a title, choose a custom color as in the previous graphs. (Bonus: use factor() with levels option to display x axis values with no observations on the plot.)
```{r}
# is.na(pas_fac)
pas_fac = as.factor(Cars93$Passengers)
plot(pas_fac, ylab="Frequency", xlab = "Passenger capacity (in persons)", main="Passenger Capacity of 93 Cars on Sale in the USA in 1993", col=15)
```

## Task 4 
a) Make a histogram of the variable `Length` (the default breaks are fine). Other aspects of the plot should be as in part (a) in Task 2.
```{r}
hist(Cars93$Length, xlab="Length (in inches)", main="Length of 93 Cars on Sale in the USA in 1993", cex.lab=1.2, cex.axis=1.2, col=4)
```

b) What is the shape of the distribution?
The distribution is approximately symmetric.

c) Find the mean and median of the variable `Length.` Do these answers meet your expectations from part (b)?
```{r}
mean(Cars93$Length)
median(Cars93$Length)
```
The mean length is 183.2043, and the median is 183. This does meet my expectations from part (b) because these two measures of central tendency are close enough to justify calling the distribution approximately symmetric.

You are done! 🏁 Don't forget to commit and push your .Rmd file to your Github repository before the due date.


